PACE 2024: OCMu64, reductions and a branch-and-bound solver for\\ One-sided Crossing Minimization

### 2024-09-03 tex/poster.tex
various changes

      \begin{block}{One-sided crossing minimizations}
        Given is a bipartite graph $(A, B)$ that is drawn in the plane at points
        $(i, 0)$ and $(j,1)$ for components $A$ and $B$ respectively. The ordering of $A$
        is fixed. The goal is the find an ordering of $B$ that minimizes the number of
        crossings when edges are drawn as straight lines.

@@

        \vspace{-1em}
        \centering
        \includegraphics[scale=2]{fig/simple.pdf}
      \end{block}

      \begin{block}{Strongly fixed pairs}
        We call $u, v \in B$ a \textbf{strongly fixed pair} if their degrees are $n$
        and $m$ and for all $0 \leq i < n$, $N(u)_i \leq N(v)_{\lfloor i\cdot m/n
          \rfloor}$.
        Intuitively, when \emph{the $x\%$'th neighbour of $u$ is left of the $x\%$'th
        neighbour of $v$}. Additionally, at least one of these inequalities must be strict.
        Note that this implies $r(u, v) < 0$.

@@

        \textbf{The lemma is optimal:}
        Without further assumptions on $(A, <)$ or on the other elements of $B$, it is not
          possible to prove $u \b v$ if $(u, v)$ is not strongly fixed.
        Suppose that $u, v \in B$ with $N(u) \neq N(v)$ are not strongly fixed, and let $i$ be
        such that $b:=N(u)_i > N(v)_{\lfloor i\cdot m/n\rfloor}=:a$. Now take sets of vertices $L, M,
          R \subseteq A$ with $a < M < b$, with $L < \min(N(v))$ and with $R > \max(N(u))$.
          Add a node $x \in B$ connected to $L, M, R$. With the correct numbers of vertices in
          $L, M, $ we can make it so that $c(v, x, u) < \min(c(x, u, v), c(u, v, x))$, so that
        $v$ has to go before $u$.

        \centering
        \includegraphics[scale=1.5]{fig/strongly-fixed.pdf}

@@

      %% \begin{block}{Weak variants}
      %%   It is also possible to consider \emph{weak} variants of the above lemmas that
      %%   only imply that $u < v$ in \emph{some} optimal solution. This requires careful
      %%   handling of cycles like $u\be v\be w\be u$.
      %% \end{block}

        \centering
        \includegraphics[scale=2]{fig/blocking-set.pdf}
    \end{column}

@@

    \begin{column}{\colwidth}
      \begin{block}{Gluing}
        A second reduction strategy \emph{glues} vertices $u$ and $v$ together.
        I.e., when $u$ and $v$ are consucutive in \emph{all} optimal solutions,
        we \textbf{glue} them together and treat them as a single vertex.

        \textbf{Gluing identical nodes:}
        When $u$ and $v$ have the same neighbours, we can glue $u$ and $v$.

        \begin{center}
          \vspace{-2em}
        \includegraphics[scale=2]{fig/gluing.pdf}
        \end{center}

@@

        \begin{description}
        \item[Graph simplification] We drop degree-$0$ vertices, merge identical
          vertices, and split the graph into \emph{independent} components.
          The initial solution uses the median heuristic and a local search that tries to move
          slices and optimally insert them. We then re-label all nodes in order to make
          memory accesses more efficient.
        \item[Fixed pairs] We find all strongly fixed pairs and store them. For the
          exact track we also find practically fixed pairs.
          For each tail we search for \emph{tail-local} practically fixed pairs. In each state, we only
          try vertices $u\in T$ not fixed by another $v\in T$.
        \item[Gluing] We use the greedy gluing strategy. Our implementation
          of practical gluing contained a bug, so we did not use this.
        \item[Tail cache] In each step, we search for the longest suffix of $T$ that
          was seen before, and reuse (the lower bound on) its score. We also
          cache the tail-local practically fixed pairs.
        \item[Optimal insert] Instead of simply appending $u$ to $P$, we
          insert it in the optimal position. The
          implementation is tricky because it interacts in complicated ways with
          the caching of results for each tail.
        \end{description}
      \end{block}

>>>>

      \begin{block}{One-sided crossing minimization}
        Given is a bipartite graph $(A, B)$ that is drawn in the plane at points
        $(i, 0)$ and $(j,1)$ for components $A$ and $B$ respectively. The ordering of $A$
        is fixed. The goal is the find an ordering of $B$ that minimizes the number of
        crossings when edges are drawn as straight lines.

@@

        \vspace{-1.5em}
        \centering
        \includegraphics[scale=2]{fig/simple.pdf}
      \end{block}

      \begin{block}{Strongly fixed pairs}
        We call $u, v \in B$ a \textbf{strongly fixed pair} if $N(u) \neq N(v)$
        and their degrees are $n$
        and $m$ and for all $0 \leq i < n$,
        $$N(u)_i \leq N(v)_{\lfloor i\cdot m/n \rfloor}.$$
        That is: for all $x$, \emph{the $x\%$'th neighbour of $u$ is left of the $x\%$'th
        neighbour of $v$}.
        %% Note that this implies $r(u, v) < 0$.

        \begin{center}
          \vspace{-2em}
        \includegraphics[scale=1.5]{fig/strongly-fixed.pdf}
          \vspace{-2em}
        \end{center}

@@

        \textbf{The lemma is optimal:}
        Without further assumptions on $(A, <)$ or on the other elements of $B$, it is not
          possible to prove $u \b v$ if $(u, v)$ is not strongly fixed.
        Suppose that $u, v \in B$ with $N(u) \neq N(v)$ are not strongly fixed, and let $i$ be
        such that $b:=N(u)_i > N(v)_{\lfloor i\cdot m/n\rfloor}=:a$. Now take sets of vertices $L, M,
          R \subseteq A$ with $a < M < b$, with $L < \min(N(v))$ and with
          $\max(N(u)) < R$.
          Add a node $x \in B$ connected to $L$, $M$, $R$. With the correct numbers of vertices in
          $L$, $M$, $R$ we can make it so that $c(v, x, u) < \min(c(x, u, v), c(u, v, x))$, so that
        $v$ has to go before $u$.

        \centering
          \vspace{-0.5em}
        \includegraphics[scale=2]{fig/blocking-set.pdf}
          \vspace{-1em}

@@

      %% \begin{block}{Weak variants}
      %%   It is also possible to consider \emph{weak} variants of the above lemmas that
      %%   only imply that $u < v$ in \emph{some} optimal solution. This requires careful
      %%   handling of cycles like $u\be v\be w\be u$.
      %% \end{block}
    \end{column}

@@

    \begin{column}{\colwidth}
      \begin{block}{Gluing}
        When $u$ and $v$ are consecutive in \emph{all} optimal solutions,
        we \textbf{glue} them together and treat them as a single vertex.

        \textbf{Identical nodes:}
        When $u$ and $v$ have the same neighbours, we glue them.

        \begin{center}
          \vspace{-1em}
        \includegraphics[scale=2]{fig/gluing.pdf}
        \end{center}

@@

        \ragnar{A figure of prefix and tail and optimal insert here would be
          nice to make the section less boring.}

        \begin{description}
        \item[Graph simplification]
          %% We drop degree-$0$ vertices, merge identical
          %% vertices, and split the graph into \emph{independent} components.
          The initial solution uses the median heuristic followed by a simple local search that tries to move
          slices and optimally insert them.
          We then re-label all nodes in order to make memory accesses more efficient.
        \item[Fixed pairs] We find all strongly fixed pairs and store them. For the
          exact track we also find practically fixed pairs.
          For each tail we search for \emph{tail-local} practically fixed pairs. In each state, we only
          try vertices $u\in T$ not fixed after another $v\in T$.
        \item[Gluing\quad] We use the greedy gluing strategy. Our implementation
          of practical gluing contained a bug, so we did not use this.
        \item[Tail cache] In each step, we search for the longest suffix of $T$ that
          was seen before, and reuse (the lower bound on) its score. We also
          cache the tail-local practically fixed pairs.
        \item[Optimal insert] Instead of simply appending $u$ to $P$, we
          insert it in the optimal position. The
          implementation is tricky because it interacts in complicated ways with
          the caching of results for each tail.
        \end{description}
      \end{block}
